Towards Explainability of Machine Learning Models in Insurance Pricing

### 2020-03-05 manuscript.tex
Deploy kasaai/explain-ml-pricing to github.com/kasaai/explain-ml-pricing.git:gh-pages

---
title: Towards Explainability of Machine Learning Models in Insurance Pricing
author:
  - name: Kevin Kuo
    email: kevin@kasa.ai
    affiliation: a
    footnote: Corresponding Author
  - name: Daniel Lupton
    email: dlupton@taylorandmulder.com
    affiliation: b
address:
  - code: a
    address: Kasa AI, 3040 78th Ave SE \#1271, Mercer Island, WA 98040, USA
  - code: b
    address: Taylor & Mulder, 10508 Rivers Bend Lane, Potomac, MD 20854, USA
abstract: |
  Machine learning methods have garnered increasing interest among actuaries in
  recent years. However, their adoption by practitioners has been limited, partly
  due to the lack of transparency of these methods, as compared to generalized 
  linear models. In this paper, we discuss the need for model interpretability in
  property & casualty insurance ratemaking, propose a framework for explaining 
  models, and present a case study to illustrate the framework.

journal: "Variance"
date: "`r Sys.Date()`"
bibliography: explain-pricing.bib
#linenumbers: true
numbersections: true
# csl: elsevier-harvard.csl
output: 
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
layout: preprint, 3p, twocolumn, letterpaper, 10pt
---

# Introduction

Risk classification for property & casualty (P&C) insurance rating has traditionally
been done with one-way, or univariate, analysis techniques. In recent years, many
insurers have moved towards using generalized linear models (GLM), a multivariate
predictive modeling technique, which addresses many shortcomings of univariate
approaches, and is currently considered the gold standard in insurance risk
classification. At the same time, machine learning (ML) techniques such as deep 
neural networks have gained popularity in many industries due to their superior
predictive performance over linear models [@lecunDeepLearning2015]. In fact, there
is a fast growing body of literature on applying ML to P&C reserving
[@kuoDeepTriangleDeep2018; @wuthrichMachineLearning2018; @gabrielliNeuralNetwork2019a;
@gabrielliNeuralNetwork2019]. However, these ML techniques, often considered to be
completely “black box”, have been less successful in gaining adoption in pricing, 
which is a regulated discipline and requires a certain amount of transparency in models.

@@

The rest of this paper is organized as follows: Section \ref{ratemaking} provides
an overview of P&C ratemaking, Section \ref{need} discusses the importance of 
interpretation, Section \ref{interpretability} discusses model interpretability 
in the context of ratemaking and proposes specific tasks for model explanation,
Section \ref{application} describes current model interpretation techniques and 
applies them to the tasks defined in the previous section, and Section 
\ref{conclusion} concludes.

# Property and Casualty Ratemaking {#ratemaking}

## History of Ratemaking

Early classification ratemaking procedures were typically univariate in nature.
For example, @lange_1966 notes that (at that time) most major lines of insurance 
used univariate methods based around the same principle: distributing an overall 
indication to territorial relativities or classification relativities based on 
the extent to which they deviated from the average experience.

@bailey_simon_1960 introduced minimum bias methods, which were expanded 
throughout the 60s, 70s, and 80s. As computing power developed, minimum bias 
began to give away to GLMs, with papers such as @brown_1988 and @mildenhall_1999 
bridging the gap between the methods.

Arguably, GLMs predate minimum bias procedures by a significant margin. The term 
was coined by @nelder_wedderburn_1972, but generalizations of least squares 
linear regression date back at least to the 1930s. Like minimum bias methods, 
GLMs did not become mainstream in actuarial science for some time. For example, 
the syllabus of basic education of the Casualty Actuarial Society (CAS) does not 
seem to include any mention of GLMs 
prior to @brown_1988 in the 1990 syllabus. From there, GLMs 
seem to have received only passing mention until 2006 with the introduction of 
@anderson_2005 to the syllabus. Beginning in 2016, the CAS introduced @goldburd_2016
to the syllabus, which offers a comprehensive guide to GLMs.

## Machine Learning in Ratemaking

Paralelling the development of GLM was the development of machine learning 
algorithms throughout the middle part of the 20th century. Detailed histories 
of machine learning may be found in sources such as @nilsson_2009 and 
@wang_raj_2017. Consistent with GLMs, machine learning was relatively unpopular 
in actuarial science until the last ten years as computing power has become cheaper
and more easily available and as machine learning software packages have obviated
the need for developing analyses from scratch each time an analysis is performed.
Due to the breadth of machine learning as a field, it is difficult to identify 
the first time it entered the CAS syllabus; however, cluster analysis (in the 
form of k-means) seems to have been first included in 2011 with @robertson_2009.
More recently, the CAS MAS-I and MAS-II exams introduced in 2018 have included 
machine learning explicitly.

Within the area of ratemaking, machine learning is still in its infancy. A 
significant portion of machine learning applications to ratemaking has been in 
the context of automobile telematics, such as @gao_2018, @gao_2018_2, @gao_2019,
@roel_2018, or @wuthrich_2017. Presumably this focus has been a result of the 
high-dimensionality and complexity of telematics data, making it a field in 
which the unique abilities of machine learning techniques give a clear 
advantage over traditional approaches.

Outside of telematics, @yang_2018 use a gradient tree-boosting approach to 
capture non-linearities that would be a challenge for GLMs. @henckaerts_2018 
make use of generalized additive models (GAM) to improve predictions of GLMs.
Many researchers, in an apparent effort to demonstrate the range of possibilities
and advantages of machine learning, have approached the topic by comparing many
different machine learning algorithms within a single study, such as in
@dugas_2003, @noll_2018, @spedicato_2018. These studies make use of such varied
techniques as regression trees, boosting machines, support vector machines, and 
neural networks.

## Ratemaking Process

@@

1. Obtain relevant policy-level data
2. Prepare data for analysis
3. Perform analysis on the data, employing desired method(s) to estimate
needed rates
4. Select final rates based on rate indications
5. Present rates to the reglator, including explanation of the steps followed to 
derive the rates
6. Answer questions from regulators regarding the method employed

@@

It should be noted that communicating results is not simply a method of passing 
regulatory muster. Generating interpretable modeling output is an important - 
even essential - facet of model checking. Actuaries are bound by relevant standards
to be able to exercise appropriate judgment in selecting risk characteristics as
part of a risk classification system per Actuarial Standard of Practice 12
("Risk Classification"). Therefore, the techniques discussed 
in this paper may be viewed from the lens of providing useful information to 
regulators, but they should also be considered as part of a thorough vetting of 
any rating model.

@@

# The Need to See Inside the Black Box {#need}

Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial 
Communications") notes that "...another actuary qualified in the same practice 
area [should be able to] make an objective appraisal of the reasonableness of 
the actuary's work as presented in the actuarial report." [@asop_41] Underlying 
this requirement is an assumption that the hypothetical other actuary qualified 
in the same practice area is adequately familiar with the relevant techniques 
employed. Although the syllabus of basic education is constantly changing, there
has at times been an assumption that all techniques and assumptions that have 
ever been a part of the syllabus of basic education needn't be explained from 
first principles in general actuarial communications, and that an actuary 
practicing in the same field should be able to make an objective appraisal of 
the results from the methods found in the syllabus.
This is notable because, beginning with the introduction
of the CAS MAS-I and MAS-II examination in July of 2018, several machine learning 
models were formally included in the syllabus of basic education. These exams 
cover a wide range of topics, such as splines, clustering algorithms, decision 
trees, boosting, and principle components analysis [@cas_syllabus_2018].

@@

1. Machine learning models can be very ad hoc compared to 
traditional statistical models.
2. Because many machine learning models do not assume an underlying probability
distribution or stochastic process, they may not admit of standard metrics for model comparison 
(e.g., it's not straightforward to calculate an AIC over a neural network). 
3. Machine learning methods are often combined into ensembles that 
may not be easily separated and that may, as a collection, cease to resemble 
a single standard version of a model.
4. Machine learning models can be "black boxes" insofar as the final form of 
response curve cannot be easily predicted and may depend heavily on the 
available data (which may not, in turn, be available to the reviewer).

@@

These issues also pose particular challenges with respect to other standards.
For instance, as discussed previously, ASOP 12 requires
actuaries to be able to exercise appropriate judgment about risk classification
systems. The recent ASOP 56 ("Models") speaks to more general concerns in all
practice areas that might make use of models. ASOP 56 requires the actuary to
"make reasonable efforts to confirm that the model structure, data, assumptions,
governance and controls, and model testing and output validation are consistent
with the intended purpose." [@asop_56] All of these efforts may be hampered if
it is not possible to peer into the black box of the model.

It should also be noted that these comments only apply within the actuarial profession. Outside of 
the actuarial profession, communication of results may be more challenging. A 
2017 survey conducted by the Casualty Actuarial and Statistical Task Force of 
the National Association of Insurance Commissioners (NAIC) found that the plurality of
responding regulators identified "Filing complexity and/or a lack of resources 
or expertise" as a key challenge that impedes their ability to review GLMs or 
other predictive models [@naic_summer_2017]. Given that machine learning 
algorithms are generally regarded as more complex than GLMs, this implies that
the challenge of communicating machine learning model results is significant.

In response to the same survey, 33 state regulators noted that it would be 
helpful or very helpful for the NAIC to develop information and tools to assist
in reviewing rate filings based on GLMs, and 34 noted that it would be helpful 
to develop similar items to assist in reviewing "Other Advanced Modeling 
Techniques." One outgrowth of this need was the development of a white paper,
@naic_white_paper, on best practices for regulatory review of predictive models.
The white paper focuses on review of GLMs, particularly with respect to private
passenger automobile and homeowners' insurance. Some of the guidance offered in
this regard is therefore not strictly applicable to the review of machine 
learning models. For example, as previously noted, p-values are not a concept
that translates well to deterministic machine learning algorithms. However,
among the guidance applicable to machine learning algorithms are the following:

* Determine the extent to which the model causes premium disruption for individual 
policyholders and how the insurer will
explain the disruption to individual consumers that inquire about it.
* Determine that individual input characteristics to a predictive model are 
related to the expected loss or expense
differences in risk. Each input characteristic should have an intuitive or 
demonstrable actual relationship to expected loss or expense.
* Determine that individual outputs from a predictive model and their associated
selected relativities are not unfairly discriminatory.

@@

# Interpretability in the Ratemaking Context {#interpretability}

@@

- Ability to explain or to present in understandable terms to a human 
[@doshi-velezRigorousScience2017];
- The degree to which an observer can understand the cause of a decision 
[@biranExplanationJustification2017]; and
- A method is interpretable if a user can correctly and efficiently predict the 
method’s results [@kimExamplesAre2016].

We motivate our discussion by considering several aspects of interpretability. 
As we proceed through the points below, we aim to arrive at a more scoped and 
relevant definition of what it means for a pricing
model to be interpretable. In the remainder of this section, we clarify a couple concepts regarding 
interpretable *classes* of models and the computational transparency of ML
models, outline frameworks for understanding the communication goals of interpretability,
then discuss a potential framework for implementing ML 
interpretability in practice.

## Not All Linear Models are Interpretable {#linear-models}

In the actuarial science literature, the GLM is probably the most oft-cited
example of an easily interpretable model. Given a set of inputs, we can easily
reason about what the output of the model is. As an illustrative example, 
consider a claim severity model with driver age, sex, and vehicle age as 
predictors; assuming a log link function and letting $Y$ denote the response, we
have

@@

Here, we can tell, for example, what the model would predict for the expected 
severity if we were to increase age by a certain amount, *all else being equal*,
because the relationship between the predictor and the response is simply 
multiplication by the coefficient $\beta_1$ and applying the inverse link
function.

Another commonly cited example of an interpretable model is a decision tree. An
illustrative example is shown in Figure \@ref(fig:tree-plot1). Here, the 
prediction is arrived at by following a sequence of if-else decisions.

```{r tree-plot1, echo = FALSE, fig.cap="A simple decision tree for loss cost prediction.", out.width="100%", fig.align='center'}
knitr::include_graphics("figures/tree-plot1.png")
```

@@

Similarly, one can see that in Figure \@ref(fig:tree-plot2), larger trees are
tough to reason about. In other words, even when working within the framework
of an "interpretable" class of models, we may still end up with something that
many would consider "black box."

```{r tree-plot2, echo = FALSE, fig.cap="A more complex decision tree. This is still much simpler than typical realistic examples.", out.width="60%", fig.align='center',  fig.env='figure*'}
knitr::include_graphics("figures/tree-plot2.png")
```

## The Machinery is Not a Secret

Another occasional misconception is that we have no visibility into how some ML
models *compute* predictions, which renders them uninterpretable. Outside of 
proprietary algorithms, all common ML models, including neural networks, 
gradient boosted trees, and random forests, are well studied and have large
bodies of literature documenting their inner workings. As an example, a fitted
feedforward neural network is simply a composition of linear transformations
followed by nonlinear activation functions. As in Equation \ref{eq:badglm}, one
can write down the mathematical equation for calculating the prediction given
some inputs, but it may be difficult for a human to reason about it. We show
later that we can still provide explanations of completely "black box" models,
but is important to note that ML model predictions are still governed by
mathematical rules, and are deterministic in most cases.

## Explanations are Contextual

@hiltonConversationalProcesses1990 proposed a framework, later interpreted by
@millerExplanationArtificial2017 in the context of ML, for understanding model explanations
as *conversations* or *social interactions.* One consequence of this identification
is that explanations need to be *relevant* to the *audience*. This framework is
consistent with ASOP 41, which formulates a similar requirement in terms of an
*intended user* of the actuarial communication. In developing,
filing, and operationalizing a pricing model, one needs to accommodate a variety
of stakeholders, each of whom has a different set of questions, assumptions,
and technical capacity. First, there are internal stakeholders at the company,
which includes management and underwriters. While some of the individuals in
this audience may be technical, they are likely less familiar with predictive
modeling techniques than the actuaries and data scientists who build the models.
Next, we have the regulators, who may have limited resources to review the
models, and will focus on a specific list of questions motivated by statute
and public policy. Finally, we have potential policyholders, who have an interest (perhaps
more so than the other parties) as they are responsible for paying the premiums.

It is interesting to note that the modelers, who are most familiar with the
models, tend to be same people designing and communicating the explanations.
This poses a challenge that @millerExplainableAI2017 call "inmates running
the asylum", where the modelers design explanations for *themselves* rather
than the intended audience. For example, they may be interested in technical
questions, such as extrapolation behavior, and shape the explanations accordingly,
which may be irrelevant to a prospective policyholder.

Another point outlined in Miller's survey [@millerExplanationArtificial2017] is
that explanations are *contrastive*. In other words, people are often interested
in not why something happened, but rather why it happened instead of something
else. For example, policyholders might not care exactly how their auto premiums
are computed, but would like to know why they are being charged more than their
coworkers who drive similar vehicles. As an extension, policyholders may want
to know what they can change in order to obtain lower premiums.

## Asking and Answering the Right Questions

@@

These questions should relate to existing guidelines, such as those described in
@naic_white_paper and outlined in Section \ref{need}, standards of practice, and
regulation, and in fact should not be specific only to ML models. By
conceptualizing a set of questions, we reduce the burden of both companies and
regulators; this is especially important for the latter, who are already
resource constrained facing increasing variety of models being filed. This
format should also be familiar to actuaries who are accustomed to adhering to
specific guidelines in, for example, ASOPs. Like the ASOPs, We envision that
these questions and guidelines will be continually updated to reflect feedback
obtained and advances in research.

@@

# Applying Model Interpretation Techniques {#application}

@@

1. What are the most important predictors in the model? Put another way, to
what extent do the predictors improve the accuracy of the model?
2. How does the predicted loss cost change, on average, as we change an input?
3. For a particular policyholder, how does each characteristic contribute to
the loss cost prediction?

The techniques we utilize to answer these questions are permutation variable
importance, partial dependence plots, and additive variable attributions, 
respectively.
In our discussion, we adopt the organization of techniques and some notation
presented in @molnar2018interpretable and @biecekPMVEE, which are comprehensive
references on the most established ML interpretation techniques.

## A Simplified View of Interpretation Techniques

>>>>

\documentclass[preprint, 3p, twocolumn, letterpaper, 10pt]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Variance} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Towards Explainability of Machine Learning Models in Insurance Pricing},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc header
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\begin{frontmatter}

  \title{Towards Explainability of Machine Learning Models in Insurance Pricing}
    \author[a]{Kevin Kuo\corref{Corresponding Author}}
   \ead{kevin@kasa.ai} 
    \author[b]{Daniel Lupton}
   \ead{dlupton@taylorandmulder.com} 
      \address[a]{Kasa AI, 3040 78th Ave SE \#1271, Mercer Island, WA 98040, USA}
    \address[b]{Taylor \& Mulder, 10508 Rivers Bend Lane, Potomac, MD 20854, USA}

  \begin{abstract}
  Machine learning methods have garnered increasing interest among actuaries in
  recent years. However, their adoption by practitioners has been limited, partly
  due to the lack of transparency of these methods, as compared to generalized
  linear models. In this paper, we discuss the need for model interpretability in
  property \& casualty insurance ratemaking, propose a framework for explaining
  models, and present a case study to illustrate the framework.
  \end{abstract}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Risk classification for property \& casualty (P\&C) insurance rating has traditionally
been done with one-way, or univariate, analysis techniques. In recent years, many
insurers have moved towards using generalized linear models (GLM), a multivariate
predictive modeling technique, which addresses many shortcomings of univariate
approaches, and is currently considered the gold standard in insurance risk
classification. At the same time, machine learning (ML) techniques such as deep
neural networks have gained popularity in many industries due to their superior
predictive performance over linear models (LeCun, Bengio, and Hinton 2015). In fact, there
is a fast growing body of literature on applying ML to P\&C reserving
(Kuo 2019; Wüthrich 2018; Gabrielli, Richman, and Wüthrich 2019; Gabrielli 2019). However, these ML techniques, often considered to be
completely ``black box'', have been less successful in gaining adoption in pricing,
which is a regulated discipline and requires a certain amount of transparency in models.

@@

The rest of this paper is organized as follows: Section \ref{ratemaking} provides
an overview of P\&C ratemaking, Section \ref{need} discusses the importance of
interpretation, Section \ref{interpretability} discusses model interpretability
in the context of ratemaking and proposes specific tasks for model explanation,
Section \ref{application} describes current model interpretation techniques and
applies them to the tasks defined in the previous section, and Section
\ref{conclusion} concludes.

\hypertarget{ratemaking}{%
\section{Property and Casualty Ratemaking}\label{ratemaking}}

\hypertarget{history-of-ratemaking}{%
\subsection{History of Ratemaking}\label{history-of-ratemaking}}

Early classification ratemaking procedures were typically univariate in nature.
For example, Lange (1966) notes that (at that time) most major lines of insurance
used univariate methods based around the same principle: distributing an overall
indication to territorial relativities or classification relativities based on
the extent to which they deviated from the average experience.

Bailey and Simon (1960) introduced minimum bias methods, which were expanded
throughout the 60s, 70s, and 80s. As computing power developed, minimum bias
began to give away to GLMs, with papers such as Brown (1988) and Mildenhall (1999)
bridging the gap between the methods.

Arguably, GLMs predate minimum bias procedures by a significant margin. The term
was coined by Nelder and Wedderburn (1972), but generalizations of least squares
linear regression date back at least to the 1930s. Like minimum bias methods,
GLMs did not become mainstream in actuarial science for some time. For example,
the syllabus of basic education of the Casualty Actuarial Society (CAS) does not
seem to include any mention of GLMs
prior to Brown (1988) in the 1990 syllabus. From there, GLMs
seem to have received only passing mention until 2006 with the introduction of
Anderson et al. (2005) to the syllabus. Beginning in 2016, the CAS introduced Goldburd, Khare, and Tevet (2016)
to the syllabus, which offers a comprehensive guide to GLMs.

\hypertarget{machine-learning-in-ratemaking}{%
\subsection{Machine Learning in Ratemaking}\label{machine-learning-in-ratemaking}}

Paralelling the development of GLM was the development of machine learning
algorithms throughout the middle part of the 20th century. Detailed histories
of machine learning may be found in sources such as Nilsson (2009) and
Wang and Raj (2017). Consistent with GLMs, machine learning was relatively unpopular
in actuarial science until the last ten years as computing power has become cheaper
and more easily available and as machine learning software packages have obviated
the need for developing analyses from scratch each time an analysis is performed.
Due to the breadth of machine learning as a field, it is difficult to identify
the first time it entered the CAS syllabus; however, cluster analysis (in the
form of k-means) seems to have been first included in 2011 with Robertson (2009).
More recently, the CAS MAS-I and MAS-II exams introduced in 2018 have included
machine learning explicitly.

Within the area of ratemaking, machine learning is still in its infancy. A
significant portion of machine learning applications to ratemaking has been in
the context of automobile telematics, such as Gao, Meng, and Wuthrich (2018), Gao and Wuthrich (2018), Gao and Wuthrich (2019),
Roel, Antonio, and Claeskens (2018), or Wuthrich (2017). Presumably this focus has been a result of the
high-dimensionality and complexity of telematics data, making it a field in
which the unique abilities of machine learning techniques give a clear
advantage over traditional approaches.

Outside of telematics, Yang, Qian, and Zou (2018) use a gradient tree-boosting approach to
capture non-linearities that would be a challenge for GLMs. Henckaerts et al. (2018)
make use of generalized additive models (GAM) to improve predictions of GLMs.
Many researchers, in an apparent effort to demonstrate the range of possibilities
and advantages of machine learning, have approached the topic by comparing many
different machine learning algorithms within a single study, such as in
Dugas et al. (2003), Noll, Salzmann, and Wuthrich (2018), Spedicato, Dutang, and Petrini (2018). These studies make use of such varied
techniques as regression trees, boosting machines, support vector machines, and
neural networks.

\hypertarget{ratemaking-process}{%
\subsection{Ratemaking Process}\label{ratemaking-process}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Obtain relevant policy-level data
\item
  Prepare data for analysis
\item
  Perform analysis on the data, employing desired method(s) to estimate
  needed rates
\item
  Select final rates based on rate indications
\item
  Present rates to the reglator, including explanation of the steps followed to
  derive the rates
\item
  Answer questions from regulators regarding the method employed
\end{enumerate}

@@

It should be noted that communicating results is not simply a method of passing
regulatory muster. Generating interpretable modeling output is an important -
even essential - facet of model checking. Actuaries are bound by relevant standards
to be able to exercise appropriate judgment in selecting risk characteristics as
part of a risk classification system per Actuarial Standard of Practice 12
(``Risk Classification''). Therefore, the techniques discussed
in this paper may be viewed from the lens of providing useful information to
regulators, but they should also be considered as part of a thorough vetting of
any rating model.

@@

\hypertarget{need}{%
\section{The Need to See Inside the Black Box}\label{need}}

Within the actuarial profession, Actuarial Standard of Practice 41 (``Actuarial
Communications'') notes that ``\ldots{}another actuary qualified in the same practice
area {[}should be able to{]} make an objective appraisal of the reasonableness of
the actuary's work as presented in the actuarial report.'' (``Actuarial Standard of Practice No. 41 - Actuarial Communications'' 2010) Underlying
this requirement is an assumption that the hypothetical other actuary qualified
in the same practice area is adequately familiar with the relevant techniques
employed. Although the syllabus of basic education is constantly changing, there
has at times been an assumption that all techniques and assumptions that have
ever been a part of the syllabus of basic education needn't be explained from
first principles in general actuarial communications, and that an actuary
practicing in the same field should be able to make an objective appraisal of
the results from the methods found in the syllabus.
This is notable because, beginning with the introduction
of the CAS MAS-I and MAS-II examination in July of 2018, several machine learning
models were formally included in the syllabus of basic education. These exams
cover a wide range of topics, such as splines, clustering algorithms, decision
trees, boosting, and principle components analysis (Casualty Actuarial Society 2018).

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Machine learning models can be very ad hoc compared to
  traditional statistical models.
\item
  Because many machine learning models do not assume an underlying probability
  distribution or stochastic process, they may not admit of standard metrics for model comparison
  (e.g., it's not straightforward to calculate an AIC over a neural network).
\item
  Machine learning methods are often combined into ensembles that
  may not be easily separated and that may, as a collection, cease to resemble
  a single standard version of a model.
\item
  Machine learning models can be ``black boxes'' insofar as the final form of
  response curve cannot be easily predicted and may depend heavily on the
  available data (which may not, in turn, be available to the reviewer).
\end{enumerate}

@@

These issues also pose particular challenges with respect to other standards.
For instance, as discussed previously, ASOP 12 requires
actuaries to be able to exercise appropriate judgment about risk classification
systems. The recent ASOP 56 (``Models'') speaks to more general concerns in all
practice areas that might make use of models. ASOP 56 requires the actuary to
``make reasonable efforts to confirm that the model structure, data, assumptions,
governance and controls, and model testing and output validation are consistent
with the intended purpose.'' (``Actuarial Standard of Practice No. 56 - Modeling'' 2019) All of these efforts may be hampered if
it is not possible to peer into the black box of the model.

It should also be noted that these comments only apply within the actuarial profession. Outside of
the actuarial profession, communication of results may be more challenging. A
2017 survey conducted by the Casualty Actuarial and Statistical Task Force of
the National Association of Insurance Commissioners (NAIC) found that the plurality of
responding regulators identified ``Filing complexity and/or a lack of resources
or expertise'' as a key challenge that impedes their ability to review GLMs or
other predictive models (National Association of Insurance Commissioners 2017). Given that machine learning
algorithms are generally regarded as more complex than GLMs, this implies that
the challenge of communicating machine learning model results is significant.

In response to the same survey, 33 state regulators noted that it would be
helpful or very helpful for the NAIC to develop information and tools to assist
in reviewing rate filings based on GLMs, and 34 noted that it would be helpful
to develop similar items to assist in reviewing ``Other Advanced Modeling
Techniques.'' One outgrowth of this need was the development of a white paper,
National Association of Insurance Commissioners (2019), on best practices for regulatory review of predictive models.
The white paper focuses on review of GLMs, particularly with respect to private
passenger automobile and homeowners' insurance. Some of the guidance offered in
this regard is therefore not strictly applicable to the review of machine
learning models. For example, as previously noted, p-values are not a concept
that translates well to deterministic machine learning algorithms. However,
among the guidance applicable to machine learning algorithms are the following:

\begin{itemize}
\tightlist
\item
  Determine the extent to which the model causes premium disruption for individual
  policyholders and how the insurer will
  explain the disruption to individual consumers that inquire about it.
\item
  Determine that individual input characteristics to a predictive model are
  related to the expected loss or expense
  differences in risk. Each input characteristic should have an intuitive or
  demonstrable actual relationship to expected loss or expense.
\item
  Determine that individual outputs from a predictive model and their associated
  selected relativities are not unfairly discriminatory.
\end{itemize}

@@

\hypertarget{interpretability}{%
\section{Interpretability in the Ratemaking Context}\label{interpretability}}

@@

\begin{itemize}
\tightlist
\item
  Ability to explain or to present in understandable terms to a human
  (Doshi-Velez and Kim 2017);
\item
  The degree to which an observer can understand the cause of a decision
  (Biran and Cotton 2017); and
\item
  A method is interpretable if a user can correctly and efficiently predict the
  method's results (Kim, Khanna, and Koyejo 2016).
\end{itemize}

We motivate our discussion by considering several aspects of interpretability.
As we proceed through the points below, we aim to arrive at a more scoped and
relevant definition of what it means for a pricing
model to be interpretable. In the remainder of this section, we clarify a couple concepts regarding
interpretable \emph{classes} of models and the computational transparency of ML
models, outline frameworks for understanding the communication goals of interpretability,
then discuss a potential framework for implementing ML
interpretability in practice.

\hypertarget{linear-models}{%
\subsection{Not All Linear Models are Interpretable}\label{linear-models}}

In the actuarial science literature, the GLM is probably the most oft-cited
example of an easily interpretable model. Given a set of inputs, we can easily
reason about what the output of the model is. As an illustrative example,
consider a claim severity model with driver age, sex, and vehicle age as
predictors; assuming a log link function and letting \(Y\) denote the response, we
have

@@

Here, we can tell, for example, what the model would predict for the expected
severity if we were to increase age by a certain amount, \emph{all else being equal},
because the relationship between the predictor and the response is simply
multiplication by the coefficient \(\beta_1\) and applying the inverse link
function.

Another commonly cited example of an interpretable model is a decision tree. An
illustrative example is shown in Figure \ref{fig:tree-plot1}. Here, the
prediction is arrived at by following a sequence of if-else decisions.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/tree-plot1}

}

\caption{A simple decision tree for loss cost prediction.}\label{fig:tree-plot1}
\end{figure}

@@

Similarly, one can see that in Figure \ref{fig:tree-plot2}, larger trees are
tough to reason about. In other words, even when working within the framework
of an ``interpretable'' class of models, we may still end up with something that
many would consider ``black box.''

\begin{figure*}

{\centering \includegraphics[width=0.6\linewidth]{figures/tree-plot2}

}

\caption{A more complex decision tree. This is still much simpler than typical realistic examples.}\label{fig:tree-plot2}
\end{figure*}

\hypertarget{the-machinery-is-not-a-secret}{%
\subsection{The Machinery is Not a Secret}\label{the-machinery-is-not-a-secret}}

Another occasional misconception is that we have no visibility into how some ML
models \emph{compute} predictions, which renders them uninterpretable. Outside of
proprietary algorithms, all common ML models, including neural networks,
gradient boosted trees, and random forests, are well studied and have large
bodies of literature documenting their inner workings. As an example, a fitted
feedforward neural network is simply a composition of linear transformations
followed by nonlinear activation functions. As in Equation \ref{eq:badglm}, one
can write down the mathematical equation for calculating the prediction given
some inputs, but it may be difficult for a human to reason about it. We show
later that we can still provide explanations of completely ``black box'' models,
but is important to note that ML model predictions are still governed by
mathematical rules, and are deterministic in most cases.

\hypertarget{explanations-are-contextual}{%
\subsection{Explanations are Contextual}\label{explanations-are-contextual}}

Hilton (1990) proposed a framework, later interpreted by
Miller (2017) in the context of ML, for understanding model explanations
as \emph{conversations} or \emph{social interactions.} One consequence of this identification
is that explanations need to be \emph{relevant} to the \emph{audience}. This framework is
consistent with ASOP 41, which formulates a similar requirement in terms of an
\emph{intended user} of the actuarial communication. In developing,
filing, and operationalizing a pricing model, one needs to accommodate a variety
of stakeholders, each of whom has a different set of questions, assumptions,
and technical capacity. First, there are internal stakeholders at the company,
which includes management and underwriters. While some of the individuals in
this audience may be technical, they are likely less familiar with predictive
modeling techniques than the actuaries and data scientists who build the models.
Next, we have the regulators, who may have limited resources to review the
models, and will focus on a specific list of questions motivated by statute
and public policy. Finally, we have potential policyholders, who have an interest (perhaps
more so than the other parties) as they are responsible for paying the premiums.

It is interesting to note that the modelers, who are most familiar with the
models, tend to be same people designing and communicating the explanations.
This poses a challenge that Miller, Howe, and Sonenberg (2017) call ``inmates running
the asylum'', where the modelers design explanations for \emph{themselves} rather
than the intended audience. For example, they may be interested in technical
questions, such as extrapolation behavior, and shape the explanations accordingly,
which may be irrelevant to a prospective policyholder.

Another point outlined in Miller's survey (Miller 2017) is
that explanations are \emph{contrastive}. In other words, people are often interested
in not why something happened, but rather why it happened instead of something
else. For example, policyholders might not care exactly how their auto premiums
are computed, but would like to know why they are being charged more than their
coworkers who drive similar vehicles. As an extension, policyholders may want
to know what they can change in order to obtain lower premiums.

\hypertarget{asking-and-answering-the-right-questions}{%
\subsection{Asking and Answering the Right Questions}\label{asking-and-answering-the-right-questions}}

@@

These questions should relate to existing guidelines, such as those described in
National Association of Insurance Commissioners (2019) and outlined in Section \ref{need}, standards of practice, and
regulation, and in fact should not be specific only to ML models. By
conceptualizing a set of questions, we reduce the burden of both companies and
regulators; this is especially important for the latter, who are already
resource constrained facing increasing variety of models being filed. This
format should also be familiar to actuaries who are accustomed to adhering to
specific guidelines in, for example, ASOPs. Like the ASOPs, We envision that
these questions and guidelines will be continually updated to reflect feedback
obtained and advances in research.

@@

\hypertarget{application}{%
\section{Applying Model Interpretation Techniques}\label{application}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  What are the most important predictors in the model? Put another way, to
  what extent do the predictors improve the accuracy of the model?
\item
  How does the predicted loss cost change, on average, as we change an input?
\item
  For a particular policyholder, how does each characteristic contribute to
  the loss cost prediction?
\end{enumerate}

The techniques we utilize to answer these questions are permutation variable
importance, partial dependence plots, and additive variable attributions,
respectively.
In our discussion, we adopt the organization of techniques and some notation
presented in Molnar and others (2018) and Biecek and Burzykowski (2019), which are comprehensive
references on the most established ML interpretation techniques.

\hypertarget{a-simplified-view-of-interpretation-techniques}{%
\subsection{A Simplified View of Interpretation Techniques}\label{a-simplified-view-of-interpretation-techniques}}
